### 2012-05-11 inst/examples/value_of_information.Rmd
added value of information example

`ro cache=TRUE, tidy=FALSE, warning=FALSE, comment=NA, message=FALSE, refresh=2 or`

@@

# Sethi Model
 * author Carl Boettiger, <cboettig@gmail.com>
 * license: CC0

 Implements a numerical version of the SDP described in `r knitcitations::citep("10.1016/j.jeem.2004.11.005")`.

Clear the workspace and load package dependencies: 
``` {r  setup, echo=FALSE}
rm(list=ls())   
require(pdgControl)
require(reshape2)
require(ggplot2)
require(data.table)
rm(list=ls())
````

@@

``` {r }
sigma_g <- 0.1    # Noise in population growth
sigma_m <- 0.1     # noise in stock assessment measurement
sigma_i <- 0.1     # noise in implementation of the quota
````

we'll use log normal noise functions. 
For Reed, only `z_g` will be random.
Sethi et al will add the other forms of noise:

``` {r  noise_dists}
z_g <- function() rlnorm(1,  0, sigma_g) # mean 1
z_m <- function() rlnorm(1,  0, sigma_m) # mean 1
z_i <- function() rlnorm(1,  0, sigma_i) # mean 1
````

@@

### Calculate the transition matrix (with noise in growth only)

@@

In the Sethi case, computing the distribution over multiple sources of noise is actually quite difficult.  Simulation turns out to be more efficient than numerically integrating over each distribution.  This code parallelizes the operation over four cores, but can be scaled to an arbitrary cluster.

``` {r  parallel}
require(snowfall) 
sfInit(parallel=TRUE, cpu=4)
SDP_Mat <- SDP_by_simulation(f, pars, x_grid, h_grid, z_g, z_m, z_i, reps=999)
````
Note that `SDP_Mat` is specified from the calculation above, as are our grids and our profit function. `OptTime` is the stopping time.  `xT` specifies a boundary condition at the stopping time. A reward for meeting this boundary must be specified for it to make any difference.  `delta` indicates the economic discount rate. Again, details are in the function documentation.

### Find the optimum by dynamic programming

Bellman's algorithm to compute the optimal solution for all possible trajectories.

@@

### Simulate

Now we'll simulate 100 replicates of this stochastic process under the optimal harvest policy determined above.

``` {r  simulate }
sims <- lapply(1:100, function(i){
  ForwardSimulate(f, pars, x_grid, h_grid, x0=K, opt$D, z_g, z_m, z_i)
})
````

The forward simulation algorithm needs an initial condition `x0` which we set equal to the carrying capacity, as well as our population dynamics `f`, parameters `pars`, grids, and noise coefficients.  Recall in the Reed case only `z_g`, growth, is stochastic, while in the Sethi example we now have three forms of stochasticity -- growth, measurement, and implementation.

@@

``` {r  tidy}
dat <- melt(sims, id=names(sims[[1]]))  
dt <- data.table(dat)
setnames(dt, "L1", "reps") # names are nice
````

@@

``` {r  onerep }
ggplot(subset(dt,reps==1)) +
  geom_line(aes(time, fishstock)) +
  geom_abline(intercept=opt$S, slope = 0) +
  geom_line(aes(time, harvest), col="darkgreen") 
````

@@

``` {r  all}
p1 <- ggplot(dt) + geom_abline(intercept=opt$S, slope = 0) 
p1 + geom_line(aes(time, fishstock, group = reps), alpha = 0.2)
````

@@

``` {r  harvestplot}
p1 + geom_line(aes(time, harvest, group = reps), alpha = 0.1, col="darkgreen")
````

@@

``` {r  escapement}
p1 + geom_line(aes(time, escapement, group = reps), alpha = 0.1, col="darkgrey")
````

### Visualizing the optimal policy

Note that when the boundary is sufficiently far away, i.e. for the first couple timesteps, the optimal policy is stationary.  The optimal policy is shown here over time, where the color indicates the harvest recommended for each possible stock value at that time (shown on the vertical axis). Observe that the solution does not correspond to a simple Reed-style rule.

``` {r  policy, fig.width=9}
policy <- melt(opt$D)
policy_zoom <- subset(policy, x_grid[Var1] < max(dt$fishstock) )
p5 <- ggplot(policy_zoom) + 
  geom_point(aes(Var2, (x_grid[Var1]), col=h_grid[value])) + 
  labs(x = "time", y = "fishstock") +
  scale_colour_gradientn(colours = rainbow(4)) +
  geom_abline(intercept=opt$S, slope = 0) 
p5
````

The harvest intensity is limited by the stock size.

``` {r  policy2, fig.width=10}
p6 <- ggplot(policy_zoom) + 
  geom_point(aes(Var2, (x_grid[Var1]), col=x_grid[Var1] - h_grid[value])) + 
  labs(x = "time", y = "fishstock") +
  scale_colour_gradientn(colours = rainbow(4)) +
  geom_abline(intercept=opt$S, slope = 0) 
p6 + geom_line(aes(time, fishstock, group = reps), alpha = 0.1, data=dt)
````

>>>>

`ro cache=TRUE, tidy=FALSE, warning=FALSE, comment=NA, message=FALSE, refresh=2 fig.width=10, fig.height=10 or`

@@

# Calculating the value of information
 * author Carl Boettiger, <cboettig@gmail.com>
 * license: CC0

 Implements a numerical version of the SDP described in `r knitcitations::citep("10.1016/j.jeem.2004.11.005")`.
 Compute the optimal solution under different forms of uncertainty and compare the results.

``` {r  setup, echo=FALSE}
rm(list=ls())   
require(pdgControl)
require(reshape2)
require(ggplot2)
require(data.table)
rm(list=ls())
````

@@

@@

# Scenarios:

@@

In the Sethi case, computing the distribution over multiple sources of noise is actually quite difficult.  Simulation turns out to be more efficient than numerically integrating over each distribution.

## No Uncertainty

``` {r noisefns}
sigma_g <- 0.0    # Noise in population growth
sigma_m <- 0.0     # noise in stock assessment measurement
sigma_i <- 0.0     # noise in implementation of the quota
z_g <- function() rlnorm(1,  0, sigma_g) # mean 1
z_m <- function() rlnorm(1,  0, sigma_m) # mean 1
z_i <- function() rlnorm(1,  0, sigma_i) # mean 1
````
Find the transition matrix

``` {r  SDP_Mat}
SDP_Mat <- determine_SDP_matrix(f, pars, x_grid, h_grid, sigma_g )
````

Find the optimum solution

@@

Simulate

``` {r  simulate }
sims_known <- lapply(1:100, function(i){
  ForwardSimulate(f, pars, x_grid, h_grid, x0=K, opt$D, z_g, z_m, z_i, profit)
})
````

## Growth uncertainty

``` {r }
sigma_g <- 0.1    # Noise in population growth
sigma_m <- 0.0     # noise in stock assessment measurement
sigma_i <- 0.0     # noise in implementation of the quota
z_g <- function() rlnorm(1,  0, sigma_g) # mean 1
z_m <- function() rlnorm(1,  0, sigma_m) # mean 1
z_i <- function() rlnorm(1,  0, sigma_i) # mean 1
````
Find the transition matrix

``` {r }
SDP_Mat <- determine_SDP_matrix(f, pars, x_grid, h_grid, sigma_g )
````

Find the optimum solution

``` {r }
opt <- find_dp_optim(SDP_Mat, x_grid, h_grid, OptTime=25, xT=0, 
                     profit, delta=0.05, reward=0)
````

Simulate

``` {r }
sims_g <- lapply(1:100, function(i){
  ForwardSimulate(f, pars, x_grid, h_grid, x0=K, opt$D, z_g, z_m, z_i, profit)
})
````

## Growth & stock measurement uncertainty

``` {r }
sigma_g <- 0.1    # Noise in population growth
sigma_m <- 0.1     # noise in stock assessment measurement
sigma_i <- 0.0     # noise in implementation of the quota
z_g <- function() rlnorm(1,  0, sigma_g) # mean 1
z_m <- function() rlnorm(1,  0, sigma_m) # mean 1
z_i <- function() rlnorm(1,  0, sigma_i) # mean 1
````

Find the transition matrix.  Use the simulation method to account for the extra uncertainties

``` {r }
require(snowfall) 
sfInit(parallel=TRUE, cpu=4)
SDP_Mat <- SDP_by_simulation(f, pars, x_grid, h_grid, z_g, z_m, z_i, reps=999)
````

Find the optimum solution

``` {r }
opt <- find_dp_optim(SDP_Mat, x_grid, h_grid, OptTime=25, xT=0, 
                     profit, delta=0.05, reward=0)
````

Simulate

``` {r }
sims_gm <- lapply(1:100, function(i){
  ForwardSimulate(f, pars, x_grid, h_grid, x0=K, opt$D, z_g, z_m, z_i, profit)
})
````

## Growth, stock measurement & implementation uncertainty

``` {r }
sigma_g <- 0.1    # Noise in population growth
sigma_m <- 0.1     # noise in stock assessment measurement
sigma_i <- 0.1     # noise in implementation of the quota
z_g <- function() rlnorm(1,  0, sigma_g) # mean 1
z_m <- function() rlnorm(1,  0, sigma_m) # mean 1
z_i <- function() rlnorm(1,  0, sigma_i) # mean 1
````

Find the transition matrix.  Use the simulation method to account for the extra uncertainties

``` {r }
require(snowfall) 
sfInit(parallel=TRUE, cpu=4)
SDP_Mat <- SDP_by_simulation(f, pars, x_grid, h_grid, z_g, z_m, z_i, reps=999)
````

Find the optimum solution

``` {r }
opt <- find_dp_optim(SDP_Mat, x_grid, h_grid, OptTime=25, xT=0, 
                     profit, delta=0.05, reward=0)
````

Simulate

``` {r }
sims_gmi <- lapply(1:100, function(i){
  ForwardSimulate(f, pars, x_grid, h_grid, x0=K, opt$D, z_g, z_m, z_i, profit)
})
````

@@

``` {r  tidy}

sims <- list(known = sims_known, growth = sims_g, growth_stock = sims_gm, growth_stock_harvest = sims_gmi)

dat <- melt(sims, id=names(sims_known[[1]]))  
dt <- data.table(dat)
setnames(dt, c("L2", "L1"), c("reps", "uncertainty")) # names are nice
````

@@

``` {r  onerep }
ggplot(subset(dt,reps==1)) +
  geom_line(aes(time, fishstock)) +
  geom_abline(intercept=opt$S, slope = 0) +
  geom_line(aes(time, harvest), col="darkgreen") + 
  facet_wrap(~uncertainty) 
````

@@

``` {r  all}
p1 <- ggplot(dt) + geom_abline(intercept=opt$S, slope = 0) 
p1 + geom_line(aes(time, fishstock, group = reps), alpha = 0.2) + facet_wrap(~uncertainty)
````

@@

``` {r  harvestplot}
p1 + geom_line(aes(time, harvest, group = reps), alpha = 0.1, col="darkgreen") + facet_wrap(~uncertainty)
````

@@

``` {r  escapement}
p1 + geom_line(aes(time, escapement, group = reps), alpha = 0.1, col="darkgrey") + facet_wrap(~uncertainty)
````

``` {r }
ggplot(subset(dt,reps==1)) +
  geom_line(aes(time, profit_fishing)) +
  geom_line(aes(time, policy_cost), col="darkblue") + facet_wrap(~uncertainty)

````

``` {r }
profits <-dt[ , sum(profit), by=c("reps", "uncertainty")] 
ggplot(profits) + geom_histogram(aes(V1)) + facet_wrap(~uncertainty)
````

Summary stats

``` {r }
profits[, mean(V1), by=uncertainty]
profits[, sd(V1), by=uncertainty]
````
